The Evolution of Phenotypically Invariant Genetic Networks

### 2017-06-07 kalman_walk.tex
added to attic

@@

\section{(Old Notes) List of papers with some notes.}

R. E. Kalman, “Mathematical description of linear dynamical systems,” Journal of the Society for Industrial and Applied Mathematics, Series A: Control, vol. 1, no. 2, pp. 152–192, 1963.
This is a good overview written by Kalman. Probably the most important paper on this list.

Minimal state-space realization in linear system theory: an overview. B. De Schutter. 2000.
Another, more recent, overview that I found helpful to understand controllability and observability.

The Inverse Problem of Stationary Covariance Generation. BDO Anderson. 1969.

Equivalence of Linear Time- Invariant Dynamical Systems. BDO Anderson, BW Necomb, RE Kalman. 1966.

B. Ho and R. E. Kalman, “Effective construction of linear state-variable models from input/output functions,” Proceedings of the 3rd Annual Allerton Conference on Circuit and System Theory, vol. 1, no. 2, p. 449459, 1965.

Dynamic Structure Functions for the Reverse Engineering of LTI Networks. J Goncalves. 2007.
I have not read this paper yet, but apparently it outlines a method for recovering the state space equation from specific sets of measurements.

New Developments in Systems Theory Relevant to Biology. RE Kalman. 1968.
Not technical but kind of interesting.

Old and New Directions of Research in Systems Theory. RE Kalman. 2010?

Minimality and Observability of Group Systems. HA Loeliger, GD Forney Jr., T Mittelholzer, MD Trott. 1994. 
Another paper I just found. It looks like it might be useful.

Partial Realization of Descriptor Systems. P Benner and VI Sokolov. 2006. 
I think this is about recovering a finite realization from infinite systems. Maybe it would be interesting to think about the similarity between transfer functions though? If selection only cares about a specific interval of the transfer function, say $G(10)$, then maybe two partial transfer functions will be effecrively equivlaent?

There are also the previous papers I sent on identifiability and distinguishability. In the original Bellman and Astrom paper where they introduce ``structural identifiability'' they show that a linear dynamical system is identifiabile if and only if it is also controllable and observable.

>>>>

\section{Algorithm for Gene Network Growth}

Let $A^(d)$ be a gene network with index $(d)$; $(d)$ is the difference between the specific realization's dimension and the dimension of a minimal realization of the requisite transfer function, such that $A^{(0)}$ is a minimal realization, $A^{(1)}$ has one superfluous dimension and $A^{(-1)}$ is one short.

Algorithm for constructing $A^{(d+1)}$ from $A^{(d)}$,

\begin{align*}
  A^{(d+1)} = \left[ 
  \begin{array}{c|c}
    A^{(d)} & 0 \\
    r_{1} \dots r_{n+d} & r_{n+d+1}
  \end{array}
  \right].
  \end{align*}

  Two systems have equivalent external dynamics iff their transfer functions are identical. Two transfer functions are identical iff $CA^{k}B = \bar{C}\bar{A}^{k}\bar{B} \qquad k = 0, 1, \dots$.

  If system $\bar{A} := A^{(1)}$ is an extension of the minimal system $A^{(0)}$ then it must be of the form,

  \begin{align*}
    \left[ \begin{array}{c|c} C & 0 \end{array} \right]
      \left[ \begin{array}{cc} V & \Psi \\ \Phi & \Gamma \end{array} \right]
        \left[ \begin{array}{cc} \Lambda & 0 \\ 0 & I \end{array} \right]
          \left[ \begin{array}{cc} V^{-1} & \Omega \\ \Theta & I \end{array} \right]
            \left[ \begin{array}{cc} B \\ \hline 0 \end{array} \right] \\
              \Phi \Theta = 0  \\
              \bar{C} \bar{A}^{k} \bar{B} = (CV\Lambda^{k}V^{-1}B + C \Psi \Theta B).
  \end{align*}

@@
